### 2014-04-05 stat246/hw1.tex
s1

Date: \today 
\vspace{1.5cm}
\subsection*{1}
Over such a thin shell of thickness $\ep$, we can assume the density function to be roughly constant. Then, evaluating the density function at any point $x$ with norm $r$ gives
\[\frac{1}{(2\pi\sigma^2)^{d/2}}\exp\left(\frac{-1}{2}x^T\sigma^{-2}Ix\right)=\frac{1}{(2\pi\sigma^2)^{d/2}}\exp\left(\frac{-r^2}{2\sigma^2}\right)\]
The hypervolume of the surface of the sphere is $S_dr^{d-1}$ (since it's a $d-1$-dimensional set), so the volume of the shell is approximately $\ep S_dr^{d-1}$. The integral of the density over the shell is approximately the volume of the shell times the value of the density, or 
\[\ep\frac{S_dr^{d-1}}{(2\pi\sigma^2)^{d/2}}\exp\left(\frac{-r^2}{2\sigma^2}\right)\]
To find the maximum of $f$, differentiate it by $r$ to obtain
\[\frac{(d-1)S_dr^{d-2}}{(2\pi\sigma^2)^{d/2}}\exp\left(\frac{-r^2}{2\sigma^2}\right)-\frac{S_dr^d}{\sigma^2(2\pi\sigma^2)^{d/2}}\exp\left(\frac{-r^2}{2\sigma^2}\right)\]
Setting this equal to $0$ and canceling gives
\[0=(d-1)-\frac{r^2}{\sigma^2}\implies r=\pm\sigma\sqrt{d-1}\]
Now, to check that this is a maximum, taking the second derivative gives us a common, positive multiplier times the following:
\[(d-1)(d-2)-\frac{r^2}{\sigma^2}(2d-1)+\frac{r^4}{\sigma^4}\]
Plugging in the previously obtained value for the stationary point, we have
\[d^2-3d+2-(d-1)(2d-1)+(d-1)^2=-2d+2\]
which becomes negative for large $d$. Thus, the point is a maximum, and we can estimate it by $\sigma\sqrt{d}$ for large $d$.
%TODO: defn of p?
\subsection{2}
By the AM-GM inequality, we have $\frac{a+b}{2}\leq\sqrt{ab}$. Since the minimum of $a$ and $b$ is bounded above by the arithmetic mean, we thus also have $\min(a,b)\leq\sqrt{ab}$.

>>>>

Date: \today 
\vspace{1.5cm}
\subsection*{1}
Over such a thin shell of thickness $\ep$, we can assume the density function to be roughly constant. Then, evaluating the density function at any point $x$ with norm $r$ gives
\[\frac{1}{(2\pi\sigma^2)^{d/2}}\exp\left(\frac{-1}{2}x^T\sigma^{-2}Ix\right)=\frac{1}{(2\pi\sigma^2)^{d/2}}\exp\left(\frac{-r^2}{2\sigma^2}\right)\]
The hypervolume of the surface of the sphere is $S_dr^{d-1}$ (since it's a $d-1$-dimensional set), so the volume of the shell is approximately $\ep S_dr^{d-1}$. The integral of the density over the shell is approximately the volume of the shell times the value of the density, or 
\[\ep\frac{S_dr^{d-1}}{(2\pi\sigma^2)^{d/2}}\exp\left(\frac{-r^2}{2\sigma^2}\right)\]
To find the maximum of $f$, differentiate it by $r$ to obtain
\[\frac{(d-1)S_dr^{d-2}}{(2\pi\sigma^2)^{d/2}}\exp\left(\frac{-r^2}{2\sigma^2}\right)-\frac{S_dr^d}{\sigma^2(2\pi\sigma^2)^{d/2}}\exp\left(\frac{-r^2}{2\sigma^2}\right)\]
Setting this equal to $0$ and canceling gives
\[0=(d-1)-\frac{r^2}{\sigma^2}\implies r=\pm\sigma\sqrt{d-1}\]
Now, to check that this is a maximum, taking the second derivative gives us a common, positive multiplier times the following:
\[(d-1)(d-2)-\frac{r^2}{\sigma^2}(2d-1)+\frac{r^4}{\sigma^4}\]
Plugging in the previously obtained value for the stationary point, we have
\[d^2-3d+2-(d-1)(2d-1)+(d-1)^2=-2d+2\]
which becomes negative for large $d$. Thus, the point is a maximum, and we can estimate it by $\sigma\sqrt{d}$ for large $d$.

Plugging $r^*+\ep$ to $f$, we get 
\[\frac{S_d(r^*+\ep)^{d-1}}{(2\pi\sigma^2)^{d/2}}\exp\left(-\frac{{r^*}^2+2r^*\ep+\ep^2}{2\sigma^2}\right)\]
Ignoring the first-order term in $\ep$ inside the exponential, we're left with
\[\frac{S_d(r^*+\ep)^{d-1}}{(2\pi\sigma^2)^{d/2}}\exp\left(-\frac{{r^*}^2}{2\sigma^2}\right)\exp\left(-\frac{\ep^2}{2\sigma^2}\right)\]
As $\ep$ grows larger, the decay from the exponential term will dominate the at-most-order-$d-1$ contribution from $(r^*+\ep)^{d-1}$, so we can approximate this as 
\[f(r^*)\exp\left(-\frac{\ep^2}{2\sigma^2}\right)\]
Thus, the value of $f$ decays as the exponential of the square of the distance from the maximum point, which means that the mass of the density in shells away from that optimal radius is in fact very low.
\subsection{2}
By the AM-GM inequality, we have $\frac{a+b}{2}\leq\sqrt{ab}$. Since the minimum of $a$ and $b$ is bounded above by the arithmetic mean, we thus also have $\min(a,b)\leq\sqrt{ab}$.
